Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-11 WSSSPE1_report/WSSSPE1_summary_report.tex
Section 3 Keynote: Merge minor changes.

\section{Keynotes\note{lead: Sou-Cheng (Terrya) Choi}} \label{sec:keynotes}

@@

Another problem he experienced was staff retention with respect to
private organizations which reward those combining research and
software expertise (the ``Google Bus''). However, he is a strong fan
of private sector sustainability sources and public-private
partnerships. He noted that making a successful business from
scientific software alone is very rare: founders overvalue while
customers undervalue. He noted that to last, an open source project
needs a minimal funding requirement even with a vibrant community --
goodwill only goes so far if you are being paid to do something else.
He talked about grant schemes of relevance in the U.S., particularly
with regard to technology transfer.~\cite{sbir-web, fased-web}

@@

His talk used an example from data reduction in Astronomy, where he
needed to remove interfering effects from the device. He built a ``bad
pixel mask'', and realized that while it was persistent, there was no
way/practice of sharing this with the data, so many researchers
repeated the same calculations. He estimated that 13 person-years were
wasted in this redundant calculation.

@@

C. Titus Brown~\cite{ged-web} argued that with regard to reusable
software, ``we should just start doing it.'' Smith noted that
documentation should be ``treated as a first class entity.''  He noted
that some software fields have excellent cultures of module reuse,
such as RubyGems~\cite{rubygems-web}, PyPI~\cite{pypi-web}, and
CPAN~\cite{cpan-web}, where there is effectively ultra-low-friction
collaboration through these repositories. This has not happened in
highly numerical, compiled language scientific software.

@@

Fernando Perez emphasized that ``open source is reproducible by
necessity'' - since reproducibility is a prerequisite for remote
collaboration.  Smith therefore defined GitHub as part of the next
stage in web development, following on from the social web of
Facebook, i.e., ``The Collaborative Web.''

>>>>

\section{Keynotes\scriptsize{\note{lead: Sou-Cheng (Terrya) Choi}, \todo{include what was presented and discussed}}} \label{sec:keynotes}

@@

Another problem he experienced was staff retention with respect to
private organizations which reward those combining research and
software expertise (the ``Google Bus''). However, he is a strong fan
of private sector sustainability sources and public-private
partnerships. He noted that making a successful business from
scientific software alone is very rare: founders overvalue while
customers undervalue. He noted that to last, an open source project
needs a minimal funding requirement even with a vibrant community --
goodwill only goes so far if you are being paid to do something else.
He talked about grant schemes of relevance in the U.S., particularly
with regard to technology transfer~\cite{sbir-web, fased-web}.

@@

Smith's talk started with an example from data reduction in Astronomy, where he
needed to remove interfering effects from the device. He built a ``bad
pixel mask'', and realized that while it was persistent, there was no
way/practice of sharing this with the data. Consequently many researchers
repeated the same calculations. He estimated that 13 person-years were
wasted in this redundant calculation.

@@

C. Titus Brown~\cite{ged-web}, a WSSSPE1 contributor and participant, argued that with regard to reusable
software, ``we should just start doing it.'' Smith noted that
documentation should be ``treated as a first class entity.''  He noted
that some software fields have excellent cultures of module reuse,
such as RubyGems~\cite{rubygems-web}, PyPI~\cite{pypi-web}, and
CPAN~\cite{cpan-web}, where there is effectively ultra-low-friction
collaboration through these repositories. This has not happened in
highly numerical, compiled language scientific software.

@@

Fernando Perez from the audience emphasized that ``open source is reproducible by
necessity'' - since reproducibility is a prerequisite for remote
collaboration.   In response, Smith defined GitHub as part of the next
stage in web development, following on from the social web of
Facebook, i.e., ``The Collaborative Web.''
